Data Analysis and Visualisation using R

### 2016-01-12 paper.tex
Larch done

The response variable {\tt Y} is continuous so regression methods are used. First, we use {\tt pairs()} to produce a matrix of pairwise scatterplots for each pair of variables ({\tt Y} and all {\tt X\textsubscript{i}}) to detect whether any observable patterns exist. {\tt X7} shows an obvious linear relationship with {\tt Y} in these plots so we use {\tt lm()} to fit a \textbf{simple linear regression} between them. The summary of this fit shows that the least squares coefficients of {\tt X7} in the linear relationship is highly significant (p-value is close to zero). Its training $R^2$ value of 0.9966 shows that the variance left unexplained by the linear fit is negligible. This fit also results in a convincing inspection plot as shown in Figure~\ref{fig:01-ash}. No obvious pattern can seen in the residuals vs fitted values plot. However, this should not rule out the possibility that {\tt Y} is also related to some other predictors. we perform a \textbf{multiple linear regression} using all available predictors and obtain an ``improved'' fit where both {\tt X3} and {\tt X7} are significant and training $R^2$ value is 0.9988. But such improvement must be taken with a grain of salt because {\em training $R^2$ values for any more flexible models (i.e. those with more predictors) must be no less than a more restrictive one}. To verify whether a simple linear model involving {\tt X7} only is the true relationship over other more flexible multiple linear models, 5-fold cross-validation has been used with best subset variables selection. As shown in Listing~\ref{ash}, cross-validation suggests $Y \sim X3+X4+X7+X8$ as the model achieving the lowest test squared error but the performance difference compared to $Y \sim X3$ is small, which is chosen for the sake of interpretability.

@@

>>>>

The response variable {\tt Y} is continuous so regression methods are used. First, we use {\tt pairs()} to produce a matrix of pairwise scatterplots for each pair of variables ({\tt Y} and all {\tt X\textsubscript{i}}) to detect whether any observable patterns exist. {\tt X7} shows an obvious linear relationship with {\tt Y} in these plots so we use {\tt lm()} to fit a \textbf{simple linear regression} between them. The summary of this fit shows that the least squares coefficients of {\tt X7} in the linear relationship are highly significant (p-value is close to zero). Its training $R^2$ value of 0.9966 shows that the variance left unexplained by the linear fit is negligible. This fit also results in a convincing inspection plot as shown in Figure~\ref{fig:01-ash}. No obvious pattern can be seen in the residuals vs fitted values plot. However, this should not rule out the possibility that {\tt Y} is also related to some other predictors. we perform a \textbf{multiple linear regression} using all available predictors and obtain an ``improved'' fit where both {\tt X3} and {\tt X7} are significant and training $R^2$ value is 0.9988. But such improvement must be taken with a grain of salt because {\em training $R^2$ values for any more flexible models (i.e. those with more predictors) must be no less than a more restrictive one}. To verify whether a simple linear model involving {\tt X7} only is the true relationship over other more flexible multiple linear models, 5-fold cross-validation has been used with best subset variables selection. As shown in Listing~\ref{ash}, cross-validation suggests $Y \sim X3+X4+X7+X8$ as the model achieving the lowest test squared error but the performance difference compared to $Y \sim X3$ is small, which is chosen for the sake of interpretability.

@@

\subsection{Larch subset}

Among predictor columns {\tt X3, X4, X5} and {\tt X9}, we observe an evident linear relationship between {\tt Y} and {\tt X5} in the pairwise scatterplots (Figure~\ref{fig:06-larch}) for every pair of variables ({\tt Y} and the predictors). A hypothesis we have about this data set is that some of the predictors are irrelevant to the response variable Y, so I decide to apply \textbf{variable selection techniques} using regsubsets() with the \textbf{best subset}, \textbf{forward stepwise} and \textbf{backward stepwise approach}. The three approaches gave exactly the same set of predictors for each number of predictors. As this data set is smaller than the previous subsets, we use 5-fold cross-validation instead of validation set approach to choose among the models. Finally, the model using all predictors (i.e. $Y \sim X3+X4+X5+X9$) achieve the lowest 5-fold cross-validated test error. The result is confirmed by a multiple linear regression fit using all these predictors where every predictor is associated with a low p-value and the training $R^2$ is as high as 0.9593. No outliers or high leverage points can be observed from the four default plots (Figure~\ref{fig:07-larch}) for the linear model.

\begin{figure}[h!]
  \includegraphics[width=\linewidth]{project/images/06-larch.png}
  \caption{Inspection plot for pairwise correlation in larch dataset.}
  \label{fig:06-larch}
\end{figure}

\begin{figure}[h!]
  \includegraphics[width=\linewidth]{project/images/07-larch.png}
  \caption{Inspection plots showing good linear fit for $\sqrt{Y} \sim X3+X4+X5+X9$ in larch dataset.}
  \label{fig:07-larch}
\end{figure}
